---
title: "eds212-day4-comp2"
format: html
editor: visual
execute: 
  warning: FALSE
---

## Exploring Data in R

```{r}
library(palmerpenguins)
library(GGally)
library(skimr)

library(tidyverse)

```

echo -- setting to false = code doesn't show when rendered

## Basic Data Exploration

-   names: get variable names

-   dim: get data dimensions

-   head: get first n lines (6 by default)

-   tail: get last n lines (6 by default)

-   summary: get basic summary statistics

```{r}
view(penguins)

```

```{r}
# check column names
names(penguins)

```

```{r}
# check the dimensions
dim(penguins)
```

```         
```

```{r}
# print the first 6 lines
# df.head() in pandas

head(penguins)
```

```{r}
# Print the last 6 lines 
tail(penguins) # See df.tail() in pandas


```

```{r}
# get summary of data, look at NAs

summary(penguins)


```

```{r}
skimr::skim(penguins)
```

tries to show a lot of types of viz to compare variables, groups

as we move along the diagonal, anytime we see a histogram its categorical??

if we look at relationship b/w bill length (mm) vs. body mass (g) (2nd column) ...?

"pairs plot"

replaces iris :)

nice version:

```{r}
# Scatterplot example 1: penguin flipper length versus body mass
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 2) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) 
```

```{r}

# Scatterplot example 2: penguin bill length versus bill depth
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 2)  +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4"))
```

```{r}
penguins %>%
  select(species, body_mass_g, ends_with("_mm")) %>% 
  GGally::ggpairs(aes(color = species)) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"))
```

## Making more exploratory charts in ggplot 2

histogram

based on variable, within aes, based on constant, outside

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm,
                            fill = island)) +
  geom_histogram() +
  theme_minimal()
```

## Box plot!

```{r}
ggplot(data = penguins, aes(x = species,
                            y = bill_length_mm)) +
  geom_boxplot(fill = "yellow", aes(color = island)) +
  geom_jitter(alpha = 0.5, aes(size = bill_depth_mm, fill = bill_depth_mm)) +
  theme_minimal()
```
